---
title: "Plotly"
output:  
  html_document:
    toc: true
    toc_float: true
---

### Introduction

A subset of the [NYC Restaurant Inspections](https://p8105.com/dataset_restaurant_inspections.html) dataset is used to create the following graphs. For the flexdashboard click [here](flexdashboard.html) 

```{r, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(plotly)
```


```{r import dataset, include = FALSE }
data(rest_inspec)

rest_inspec

#397584 observations and 18 variables 
#variables include: action, boro, building, camis, critical_flag, cuisine_description, dba, inspection_date, inspection_type, phone, record_date, score, street, violation_code, violation_description, zipcode, grade, grade_date
```

```{r explore data, include = FALSE}

distinct(rest_inspec, action) #5, "Violations were cited in the following area(s)" included in two categories, also have NA. Basically: closed, re-closed, re-open, no violation, violation cited., no NAs

distinct(rest_inspec, boro) #all 5 boroughs, with "missing" entries as well, no NAs (just "missing")

distinct(rest_inspec, critical_flag)  #3 levels: not critical, critical, not applicable, no NAs

distinct(rest_inspec, cuisine_description) #85 types, types can be a combination of different things, one misspled? "Café/Coffee/Tea", no NAs

distinct(rest_inspec, inspection_date) #from year 1900 to 2017, inspection dates become more often later on. 

distinct(rest_inspec, record_date) #one date (10/19/17) with 3 different times that varried by the second

distinct(rest_inspec, score) #from -2 to 151, integers, there is NA
distinct(rest_inspec, street) #there's NA and special characters

distinct(rest_inspec, violation_code) #99 different codes, code name includes 2 digit number and one letter, there is NA
distinct(rest_inspec, violation_description) #there is NA, might not need this

distinct(rest_inspec, zipcode) #234 zip codes, there is also NA

distinct(rest_inspec, grade) #A, B, C, P, Z, "Not yet graded", and NA

distinct(rest_inspec, grade_date) #there is NA, 2012-2017
```

### Dataset Used

* Excludes: 
  * missing borough (`boro`) information
  * observations with NAs for `score` variable.

* Includes:
  * inspection dates in 2013 to 2017 (5 year data)
  * only observations that received grades A, B, or C.
  
  Resulting dataset has 188189 observations and 19 variables. 

```{r sample, include = FALSE}
#obtain smaller sample size from dataset. 

rest_inspec_sample = rest_inspec |>
  separate(inspection_date, into = c("insp_year", "insp_month", "insp_date"), sep = "-") |> 
  filter(
    boro != "Missing", 
    insp_year %in% c("2013", "2014", "2015", "2016", "2017"), #sample from 2013-2017. 5 year data.
    grade %in% c("A", "B", "C")) |> #keep only those with A, B, or C
  select(-c(building, phone, record_date)) |>
  separate(grade_date, into = c("gr_year", "gr_month", "gr_date"), sep = "-") |>
  mutate(
    boro = as.factor(boro),
    critical_flag = as.factor(critical_flag),
    action = as.factor(action),
    cuisine_description = as.factor(cuisine_description),
    grade = as.factor(grade))

#188191 observations x 17 variables

distinct(rest_inspec_sample, boro) #all 5 boroughs. No NAs.
distinct(rest_inspec_sample, action) #3 categories. No NAs.
count(rest_inspec_sample, critical_flag) #3 levels: not critical, critical, not applicable. No NAs
count(rest_inspec_sample, cuisine_description) |> view() #85 types

#19739 restaurant names but 24713 different restaurant identifiers. Maybe because franchise like Dunkin' Donuts, Starbucks, etc.
count(rest_inspec_sample, dba) #19739 restaurant names
count(rest_inspec_sample, camis) #some restaurants have been inspected multiple times. max. of 36X, min. of 1x. 24,713 different entities of restaurants.
count(rest_inspec_sample, inspection_type) #9 types
count(rest_inspec_sample, score) #-1 to 86. 2 NAs.
count(rest_inspec_sample, violation_code) #68 types
count(rest_inspec_sample, violation_description) #64 types, also has NA.
count(rest_inspec_sample, zipcode) #231 different zipcodes
```


```{r, include = FALSE}
rest_inspec_sample1 = rest_inspec_sample |>
  drop_na(score) #drop 2 observations with NA
```


### Bar Graph: Counts Across 5 Boroughs
```{r bar, echo = FALSE}
bar = rest_inspec_sample1 |> 
  count(boro) |> 
  mutate(boro = fct_reorder(boro, n)) |> 
  plot_ly(x = ~boro, y = ~n, color = ~boro, type = "bar", colors = "viridis")

bar
```

In this dataset, Manhattan has the highest count (74.8k) while Staten Island has the least (6,355).

### Manhattan Bar Graph: Grade vs. Score

```{r boxplot, echo = FALSE}
box_ggplot = 
  rest_inspec_sample1 |> 
  filter(boro == "MANHATTAN") |>
  mutate(grade = fct_reorder(grade, score)) |> 
  ggplot(aes(x = grade, y = score, fill = boro)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplotly(box_ggplot)
```

* In Manhattan, entities that received a C grade have a median score of 34 with an outlier scoring 83 points.
  * As the grade increases from A to C, the median scores also increases.


### Manhattan Scatterplot: Zip code vs. Score

This scatterplot contains observations in Manhattan entities that sell American food and received a grade in 2017. There are 4,843 observations. 

```{r scatter, echo = FALSE}
scatter_ggplot = rest_inspec_sample1 |>
  filter(cuisine_description == "American", boro == "MANHATTAN", gr_year == 2017) |>
  ggplot(aes(x = zipcode, y = score, color = boro)) +
  geom_point(alpha = 0.25) +
  coord_cartesian()

ggplotly(scatter_ggplot)
```

* There seems to be higher concentration of entities that were scored between zip codes 10000 and 10044. The scores is more concentrated between 2 and 13.